---
title: "Project 3 - Statistics and Visualization"
author: "Ji-Eun Park"
date: "11/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
path <- "/Users/jieun/Downloads/datasci611-gh-pages-4/data/project2_2019"
files <- list.files(path)

for (i in 1:length(files)) {
  name <- substr(files[i],1,nchar(files[i])-4)
  assign(name, data.table::fread(file=paste0(path,"/",files[i])))
}
rm(i,name,path,files)
```

## Choose Variables

```{r}
datasets <- ls()

varnames <- function(i) {
  get(datasets[i]) %>% colnames()
}

lapply(1:16, function(x) varnames(x)) %>% unlist() %>% unique()

```

```{r}
a = read.csv("../data/client2.csv")[,-1]
b = read.csv("../data/incomeEntry2.csv")[,-1]
c = read.csv("../data/incomeExit2.csv")[,-1]
d = read.csv("../data/HIEntry2.csv")[,-1]
e = read.csv("../data/HIExit2.csv")[,-1]
f = read.csv("../data/DisEntry2.csv")[,-1]
g = read.csv("../data/DisExit2.csv")[,-1]

dat <- left_join(a,b)
dat <- left_join(dat,c) %>% select(-index)
dat <- left_join(dat,d) %>% select(-index)
dat <- left_join(dat,e) %>% select(-index)
dat <- left_join(dat,f) %>% select(-index)
```

### Background
**Urban Ministries of Durham (UMD)** is a non-profit organization offering food, shelter and other services necessary for people in need of those aids. This project is designed to bring meaningful insights from the supply data gathered since around 1990's which is almost since the foundation of this organization (1983). 

### Project Goal
The aim of this project is to provide UMD with useful and invaluable information about the organization and the clients and provide analytic information and suggest insightful ideas for UMD.

### Data
UMD has provided 16 datasets including such as client/provider information, disability at enty/exit, health insurance at entry/exit, income at entry/exit, noncash support received entry/exit, etc. 

### Questions of Interest
**Questions may change in final report**
1. Determine the demographics of UMD clients. What is the distribution of the clients by race, age groups, gender, race, etc?

2. How did the income and non-cash support change from entry to exit? Were there any improvements?

3. What are the status of the clients at exit? Were they in better condition (e.g. destination)? What are the reasons why the clients left?

